Realizing temporal graphs from fastest travel times

### 2023-09-20 STACS2023/nina_shortversion.tex
trying to make a shorter version for STACS

\usepackage{graphicx}
\graphicspath{{Figures/}}

@@

We only give an informal and very high-level description of the main ideas of the algorithm
The detailed description and correctness proof of the algorithm are deferred to the appendix, due to space constraints. It is well-known that the number of paths between two vertices in a static graph can be upper-bounded by a function of the feedback edge number of the graph. Intuitively, this would allow us to ``guess'' paths between a small number of vertex pairs in $G$ (the underlying graph of our instance \deltaExact). These paths would then be used to create (representatives of) fastest temporal paths among vertex pairs in $(G,\lambda)$. We would then try to use the polynomial algorithm for trees (which includes paths) to figure out the labeling $\lambda$ of the edges used by these path. An idea like this has been used for several other temporal graph problems related to the connectivity between two vertices~\cite{casteigts2021finding,FMNR22a,EMM22}.

@@

 \begin{comment}
	\begin{figure}[t]
    \centering
	\begin{tikzpicture}[xscale=2,yscale=0.4]
	%%%S_uv
	\node[vert,label=below:$u$] (v1) at (1,0) {};
	\node[vert] (v2) at (2,0) {};
	\node[vert,label=below:$w$] (v3) at (3,0) {};
	\node[vert] (v4) at (4,0) {};
	\node[vert,label=below:$v$] (v5) at (5,0) {};
	\draw (v1) -- node[timelabel] {$1$} (v2) -- node[timelabel] {$3$}  (v3) -- node[timelabel] {$5$} (v4) -- node[timelabel] {$7$} (v5);
	\draw (v1) to [out=80,in=100,distance=2cm] node[timelabel] {$9$} (v3);

@@

	\draw[red]
	(v1) edge[diredge2] [out=85,in=95,distance=2.1cm] node[pos=0.3,yshift=2,label=above:$P_{u,w}$] {} (v3) ;
\end{tikzpicture}
	\caption{An example of a temporal graph (with $\Delta \geq 9$), where the fastest temporal path $P_{u,v}$ (in blue) from $u$ to $v$ is of duration $7$,
		while the fastest temporal path $P_{u,w}$ (in red) from $u$ to a vertex $w$, that is on a path $P_{u,v}$, is of duration $1$ and is not a subpath of $P_{u,v}$.		
		\label{fig:ftpExample}}
	\end{figure}
\end{comment}

To overcome the above mentioned difficulties, we propose the following high-level strategy.
\begin{itemize}
    \item We identify a small number of ``interesting vertices'' between which we guess representative fastest paths.
    \item We show that we can determine most labels along each path up to an additive constant. We show that the number of labels we cannot determine along such a path is small. 
    \item For each additive constant and each non-determined label along such a path we introduce a variable for an Integer Linear Program (ILP). 
    \item We use constraints in the ILP to ensure that alternative connections (not using the guessed fastest path) between vertex pairs are not too fast.
    \item We compute a solution to the ILP and construct a labeling from that solution. Since the ILP has few variables and not too many constraints, this computation can be done efficiently using Lenstra's Theorem~\cite{Lenstra1983Integer}.
\end{itemize}

@@

For the remainder of this section, let $D$ be the input matrix of \deltaExact\, and let $G$ be its underlying graph, on $n$ vertices and $m$ edges.
%With $F$ we denote a minimum feedback edge set of $G$, and with $k$ the feedback edge number of $G$.
We are now ready to present our FPT algorithm. For an easier readability we split the description and analysis of the algorithm in five subsections.
We start with a preprocessing procedure for $G$, where we define a set of interesting vertices which then allows us to guess the desired structures.
Next we introduce some extra properties of our problem, that we then use to create ILP instances and their constraints.
At the end we present how to solve all instances and produce the desired labeling $\lambda$ of $G$, if possible.

\subparagraph{Preprocessing of the input.}
From $G$ we create $G'$ by
iteratively removing vertices of degree one. With $Z$ we denote the set of all removed vertices.
%by finding a spanning tree $T$ of $G'$ and set $F = E(G) \setminus E(T)$. 
%Note that $F$ is also a minimum feedback edge set of $G$.
%Lastly, we determine sets $U$, of \emph{vertices of interest}, and $U^*$ of the neighbors of vertices of interest, in the following way.
Let $T$ be a spanning tree of $G'$, that gives rise to a minimum feedback edge set $F$ (denote $k = |F|$), 
and
let $V_1 \subset V(G')$ be the set of leaves in $T$, $V_2 \subset V(G')$ the set of vertices of degree two in $T$, that are incident to at least one edge in $F$, 
and let $V_3 \subset V(G')$ be the set of vertices of degree at least $3$ in $T$. 
Then $|V_1| + |V_2| \leq 2k$, since every leaf in $T$ and every vertex in $V_2$ is incident to at least one edge in $F$,
and $|V_3| \leq |V_1|$ by the properties of trees.
We denote with 
$
U = V_1 \cup V_2 \cup V_3
$
the set of \emph{vertices of interest}. It follows that $|U| \leq 4k$.
We set $U^*$ to be the set of vertices in $V(G') \setminus U$ that are neighbors of vertices in $U$, \ie 
$
U^* = \{v \in V(G') \setminus U  :  u \in U, v \in N(u)\}.
$
Again, using the tree structure, we get that for any $u \in U$ its neighborhood is of size $|N(u)| \in O(k)$, therefore $|U^*| \in O(k^2)$.
Note that determining all of the above sets takes linear time.

\subparagraph{Guessing necessary structures. \label{sec:FPT-guessing}}
Once the sets $U, U^*$ %and $Z^*$ 
are determined, we are ready to start guessing  the necessary structures.
Note that whenever we say that we guess the fastest temporal path between two vertices, we mean that we guess the underlying path of a representative fastest temporal path between those two vertices.
In the case when we want to guess the fastest path from $u$, that (somehow) passes through the vertex $x$, and goes to $v$ via an edge (or a unique path), we write it as a fastest path of the form $u \leadsto x \rightarrow v$.
%Similarly it holds if we want to include more vertices that have to be visited by the path. 
%If there is an edge (\ie a unique fastest path) among two vertices, we denote it by $\rightarrow$, if the fastest path among two vertices is not uniquely determined, we denote it by $\leadsto$.
We guess $11$ different structures $G-1,G-2,\dots,G-11$, we present here just $G-1,G-4$, for all the others we refer the reader to the Appendix.
\begin{enumerate}[G-1.]
    \item \label{FPT-guessFTPamongU}
    The fastest temporal paths among all pairs of vertices of $U$.
    For a pair $u,v$ of vertices in $U$, there are $k!$ possible paths in $G'$ between them. 
    Therefore, we have to try all $O(k^k)$ possible paths, where at least one of them will be a fastest temporal path, respecting the value from $D$.
    %If this does not happen we conclude that $D$ is a \textsc{No} instance of \deltaExact.
    Repeating this procedure for all pairs $u,v \in U$ we get $O((k^k)^{k^2})=O(k^{k^3})$ variations of the fastest temporal paths among all vertices in $U$.

    %\item \label{FPT-guessFTPamongZstar}
    %The fastest temporal paths among all vertices in $Z^*$, which by similar arguing as for vertices in $U$, gives us $O(k^{k^3})$ guesses.

    %\item \label{FPT-guessFTPamongUstar}
    %The fastest temporal paths among all vertices in $U^*$. This gives us $O(k^{k^6})$ guesses.

    \setcounter{enumi}{3}

    \item \label{FPT-guessFTPamongUandUstar}
    The fastest temporal paths from vertices of $U$ to vertices in $U^*$,
    and the fastest temporal paths from vertices in $U^*$ to vertices in $U$.
    This gives us $O(k^{k^4})$ guesses.
\end{enumerate}
%
We create all of the guesses independently.
We start with the first one, that results in $O(k^{k^3})$ different possibilities, then we split each one of these guesses into $O(k^{k^4})$ new ones, that respond to the guessing in the next step, etc.
After creating all of the guesses we end up with $f(k)$ different cases (where $f$ is a double exponential function).
We now create one ILP instance for each case.
From now on we assume that we know exactly the structure of all paths we have guessed.

\subparagraph{Properties of \deltaExact.}

In this section we study the properties of our problem, that then help us creating constraints of our ILP instances.
	We want to determine labeling $\lambda$ of each edge of $G$.
	We start with an empty labeling of edges and try to specify each one of them.
	Note, that this does not necessarily mean that we assign numbers to the labels, but we might specify labels as variables or functions of other labels.
    We say that the label of an edge $f$ is \emph{determined with respect to} the label of the edge edge $e$, if we have determined $\lambda(f)$ as a function of $\lambda(e)$. 
	%In fact at the beginning the values will be determined with respect to each other (\ie as a function of each other).
	%In our analysis we exploit the following results greatly.
    We are now ready, to present some basic properties of our problem.
\begin{definition}
    Let $U \subseteq V(G')$ be a set of vertices of interest and let $u,v \in U$.
    A path $P = (u=v_1,v_2, \dots, v_p = v)$  in $G'$, of length at least $2$, where $v_i \notin U$ for $i \in \{ 2, 3, \dots, p-1\}$,
    is called a \emph{segment} from $u$ to $v$. We denote it as $S_{u,v}$.
    Note $S_{u,v} \neq S_{v,u}$.
\end{definition}
%From the definition it follows, that there are $O(k^2)$ segments in $G'$. \todo{True but does not follow directly just from the definition}
Observe that a temporal path in $G'$ between two vertices of interest is either a segment, or consists of a sequence of some segments. Furthermore, since we have at most $4k$ interesting vertices in $G'$, we can deduce the following important result.
\begin{corollary}\label{obs:FPT-k2segments}
    There are at most $O(k^2)$ segments in $G'$.
\end{corollary}

	\begin{observation}[$\star$]\label{obs:FirstLabelAndDuration}
		Let $P$ be the underlying path of a fastest temporal path from $u$ to $v$, where $e_1, e_p \in P$ are its first and last edge, respectively.
		Then, knowing the label $\lambda (e_1)$ of the first edge and the duration $d(P,\lambda)$ of the temporal path $(P,\lambda)$, we can uniquely determine the label $\lambda (e_p)$ of the last edge of $P$.
		Symmetrically, knowing $\lambda(e_p)$ and $d(P,\lambda)$, we can uniquely determine $\lambda(e_1)$.
	\end{observation}
 If $\lambda(e_1)$ and $\lambda(e_p)$ (from the above observation) are both unknown, then we can determine one with respect to the other.
 %\begin{definition}\label{FPT:def-segment}
%	Let $U \subseteq V(G')$ be a set of vertices of interest and let $u,v \in U$.
%	A path $P = (u=v_1,v_2, \dots, v_p = v)$  in graph $G'$, where all inner vertices are not in $U$, \ie $v_i \notin U$ for $i \in \{ 2, 3, \dots, p-1\}$,
%	is called a \emph{segment} from $u$ to $v$. We denote it as $S_{u,v}$.
%	Note $S_{u,v} \neq S_{v,u}$.
%\end{definition}

	\begin{lemma}[$\star$]\label{lemma:FPT-uv-LabelAlmostalledges}
		Let $u, v \in U$ be two arbitrary vertices of interest and suppose that $P = (u=v_1, v_2, \dots, v_p = v)$, where $p \geq 2$, 
		is a path in $G'$, which is also the underlying path of a fastest temporal path from $u$ to $v$.
		Let $\ell_{u,v} \geq 1$ be the number of vertices of interest in $P$ different to $u,v$, namely $\ell_{u,v} = |\{P \setminus \{u,v \} \} \cap U |$.
		We can determine the labeling $\lambda$ of all but $\ell_{u,v}$ edges of $P$, with respect to the label $\lambda(uv_2)$ of the first edge,
		such that the labeling $\lambda$ respects the values from $D$.
		%    Moreover, for the labels of edges that cannot be determined precisely, we introduce some extra restrictions.
	\end{lemma}

\subparagraph{Adding constraints and variables to the ILP.}
%Using the above guessed structures, and certain properties of the problem, we proceed with adding constraints to each of our ILP instances.

%Recall that a labeling $\lambda$ of $G$ satisfies $D$ if the duration of a fastest temporal path from vertex $v_i$ to $v_j$ equals $D_{v_i, v_j}$.
In order to find a labeling $\lambda$ satisfying $D$ 
%that satisfies this property 
we split our analysis in nine cases.
We consider fastest temporal paths 
where the starting vertex $v_i$ is in one of the sets $U,V(G') \setminus U, Z$,
and similarly the destination vertex $v_j$ is in one of the sets $U,V(G') \setminus U, Z$.
In each of these cases we guess the underlying path $P$ that at least one fastest temporal path from\ $v_i$ to $v_j$ follows, 
which results in one equality constraint for the labels on the path $P$. 
For all other temporal paths from $v_i$ to $v_j$ we know that they cannot be faster, so we introduce inequality constraints for them.
Note that we have to do this while keeping the total number of variables upper-bounded by some function in $k$.
To ensure this 
we split the graph into $O(k^2)$ segments, and introduce at most $3$ variables for each segment,
therefore each of our ILP instances admits $O(k^3)$ number of variables (for details see Appendix).
%
%
We now present one example of how to determine the fastest temporal paths among a fixed pair of vertices, for all other cases we refer the reader to the Appendix.

\emph{Fastest paths from $x$ to $u$, where $x \in V(G') \setminus U$ and $u \in U$.}
\begin{figure}[t]
		\centering
		%\resizebox{\linewidth}{!}{

			\begin{tikzpicture}[yscale=0.25]
				%%%S_uv
				\node[vert,label=above:$w {=} z_1$] (v1) at (1,0) {};
				\node[vert,label=above:$z_2$] (v2) at (2,0) {};
				\node[vert,label=above:$z_{r-1}$] (vp1) at (6,0) {};
				\node[vert,label=above:$z_r {=} z$] (vp) at (7,0) {};
				\draw (v1) -- (v2) -- (vp1) -- (vp);

				%%%% node w
				\node[vert,label=below:$u$] (w) at (4,-3) {};

				\draw [blue]%[transform canvas={yshift=0.5mm},thick] 
				(v1) edge[dirwave] [out=290,in=160] (w) 
				(v2) edge[diredge2] (v1);

				\draw [red]%[transform canvas={yshift=0.5mm},thick] 
				(vp) edge[dirwave] [out=240,in=20] (w) 
				(vp1) edge[diredge2] (vp);

			\end{tikzpicture}%}
		\caption{Example of a guess G-8, where we guessed the fastest paths of the form $z_2 \rightarrow w \leadsto u$ (in blue) and $z_{r-1} \rightarrow z \leadsto u$ (in red), where $u$ is a vertex of interest and $S_{w,z} = (w=z_1,z_2, \dots, z_r = z)$. \label{fig:FPT-4casesGuess}}
\end{figure}

%We want to determine the fastest path from $x$ to $u$.
%We use the similar approach as in the case~\ref{item:FPT-uToxcase2} from above.
Let $x$ be a vertex in the segment $S_{w,z} = (w=z_1,z_2, \dots, z_r = z)$.
If $S_{w,z}$ is of length $3$ or less, then we already know the fastest temporal path from every vertex in the segment to $u$ 

>>>>

\usepackage{graphicx}
\graphicspath{{../Figures/}}

@@

We first give an informal and very high-level description of the main ideas of the algorithm.
We then present the first part in more details and deffer the reader to the appendix, for the remaining part and correctness proof.
It is well-known that the number of paths between two vertices in a static graph can be upper-bounded by a function of the feedback edge number of the graph. Intuitively, this would allow us to ``guess'' paths between a small number of vertex pairs in $G$ (the underlying graph of our instance \deltaExact). These paths would then be used to create (representatives of) fastest temporal paths among vertex pairs in $(G,\lambda)$. We would then try to use the polynomial algorithm for trees (which includes paths) to figure out the labeling $\lambda$ of the edges used by these path. An idea like this has been used for several other temporal graph problems related to the connectivity between two vertices~\cite{casteigts2021finding,FMNR22a,EMM22}.

@@

	\begin{figure}[t]
    \centering
	\begin{tikzpicture}[xscale=2,yscale=0.4]
	%%%S_uv
	\node[vert,label=below:$u$] (v1) at (1,0) {};
	\node[vert] (v2) at (2,0) {};
	\node[vert,label=below:$w$] (v3) at (3,0) {};
	\node[vert] (v4) at (4,0) {};
	\node[vert,label=below:$v$] (v5) at (5,0) {};
	\draw (v1) -- node[timelabel] {$1$} (v2) -- node[timelabel] {$3$}  (v3) -- node[timelabel] {$5$} (v4) -- node[timelabel] {$7$} (v5);
	\draw (v1) to [out=80,in=100,distance=2cm] node[timelabel] {$9$} (v3);

@@

	\draw[red]
	(v1) edge[diredge2] [out=85,in=95,distance=2.1cm] node[pos=0.3,yshift=2,label=above:$P_{u,w}$] {} (v3) ;
\end{tikzpicture}
	\caption{An example of a temporal graph (with $\Delta \geq 9$), where the fastest temporal path $P_{u,v}$ (in blue) from $u$ to $v$ is of duration $7$,
		while the fastest temporal path $P_{u,w}$ (in red) from $u$ to a vertex $w$, that is on a path $P_{u,v}$, is of duration $1$ and is not a subpath of $P_{u,v}$.		
		\label{fig:ftpExample}}
	\end{figure}

To overcome the above mentioned difficulties, we propose the following high-level strategy.
\begin{itemize}
    \item We identify a small number of ``interesting vertices'' between which we guess representative fastest paths.
    \item We show that we can determine most labels along each path up to an additive constant. We show that the number of labels we cannot determine along such a path is small. 
    \item For each additive constant and each non-determined label along such a path we introduce a variable for an Integer Linear Program (ILP).
    \item We use constraints in the ILP to ensure that alternative connections (not using the guessed fastest path) between vertex pairs are not too fast.
    \item We compute a solution to each of the ILPs (if it exists) and construct a labeling from that solution. Since the ILP has few variables and not too many constraints, this computation can be done efficiently using Lenstra's Theorem~\cite{Lenstra1983Integer}.
\end{itemize}
It is important to point out that each guess of fastest paths, gives rise to one ILP. Since there are many guesses ($f(k)$ many, where $f$ is a double exponential function), we get many ILPs.
If none of the ILP instances gives a positive solution, then there exists no labeling realizing given matrix.
This is true because considering all guesses of fastest temporal paths is equivalent to exhaustively searching through all possible temporal paths between vertices.

\subsubsection{First part of the FPT algorithm}
We now present the first part of our FPT algorithm.
Let $D$ be the input matrix of \deltaExact\, and let $G$ be its underlying graph, on $n$ vertices and $m$ edges.
%\subsubsection*{Preprocessing of the input}
From the underlying graph $G$ of $D$ we first create a graph $G'$ by
iteratively removing vertices of degree one from $G$,
and denote with 
$
Z = V(G) \setminus V(G')
$, the set of removed vertices.
%Then we determine a minimum feedback edge set $F$ of $G'$.
%by finding a spanning tree $T$ of $G'$ and set $F = E(G) \setminus E(T)$. 
%Note that $F$ is also a minimum feedback edge set of $G$.
Then we determine sets $U$, of \emph{vertices of interest}, and $U^*$ of the neighbors of vertices of interest, in the following way.
Let $T$ be a spanning tree of $G'$, with $F$ being the corresponding feedback edge set of $G'$.
Let $V_1 \subset V(G')$ be the set of leaves in the spanning tree $T$, $V_2 \subset V(G')$ be the set of vertices of degree two in $T$, that are incident to at least one edge in $F$, 
and let $V_3 \subset V(G')$ be the set of vertices of degree at least $3$ in $T$. 
Then $|V_1| + |V_2| \leq 2k$, since every leaf in $T$ and every vertex in $V_2$ is incident to at least one edge in $F$,
and $|V_3| \leq |V_1|$ by the properties of trees.
We denote with 
\[U = V_1 \cup V_2 \cup V_3\]
the set of vertices of interest. It follows that $|U| \leq 4k$.
We set $U^*$ to be the set of vertices in $V(G') \setminus U$ that are neighbors of vertices in $U$, \ie 
\[U^* = \{v \in V(G') \setminus U  :  u \in U, v \in N(u)\}.\]
Again, using the tree structure, we get that for any $u \in U$ its neighborhood is of size $|N(u)| \in O(k)$, since every neighbor of $u$ is the first vertex of a (unique) path to another vertex in $U$.
It follows that $|U^*| \in O(k^2)$.
From the construction of $Z$ (by iteratively removing vertices of degree one from $G$)
it follows that $Z$ consists of disjoint trees $T_1, T_2, \dots$.
For a tree $T_i$ we denote with $u_i$ the vertex in $G'$ that is a neighbor of a vertex in $T_i$, 
and call it a \emph{clip vertex of the tree $T_i$}.
It follows that there can be many different trees $T_i$ that are incident to the same clip vertex $u_i \in V(G')$,
but each tree $T_i$ is incident to exactly one clip vertex $u_i \in V(G')$.
Since $u_i$ is the only vertex connecting all of the trees $T_i$ incident to it,
from now on we assume that a tree $T_{u_i}$ in $Z$ is a union of trees on vertices from $V(G) \setminus V(G')$,
that are clipped at the same vertex $u_i \in V(G')$.
For each of the trees $T_{u_i}$ in $Z$, we select one vertex $r_i$, that is a neighbor of the clip vertex $u_i$,
and call it \emph{a representative vertex of the tree $T_{u_i}$}.
We now define as $Z^*$ the set of representatives $r_i$ of trees $T_i \in Z$,
where the clip vertex $v_i$ of $T_i$ is a vertex of interest, \ie
\[
Z^* = \{ r_i  :  r_i \in T_i, \text{ where $T_i \in Z$, the clip vertex $u_i$ of $T_i$ is in $U$, and $r_i u_i \in E(G)$}\}.
\]
Since there are $O(k)$ vertices of interest, we get that $|Z^*| \in O(k)$.
Note that determining sets $U$, $U^*$, and $Z^*$ takes linear time.

%Let us remember the following basic property of the problem \deltaExact.
Recall that a labeling $\lambda$ of $G$ satisfies $D$ if the duration of a fastest temporal path from vertex $v_i$ to $v_j$ equals $D_{v_i, v_j}$.
%, and no other temporal path from $v_i$ to $v_j$ is faster.
In order to find a labeling that satisfies this property we split our analysis in nine cases.
We consider fastest temporal paths 
where $v_i$ is in one of the sets $U,V(G') \setminus U, Z$,
and $v_j$ is in one of the sets $U,V(G') \setminus U, Z$.
In each of these cases we guess the underlying path $P$ that at least one fastest temporal path from $v_i$ to $v_j$ follows, 
which results in one equality constraint for the labels on the path $P$. 
For all other temporal paths from $v_i$ to $v_j$ we know that they cannot be faster, so we introduce inequality constraints for them.
This results in producing $poly(n,m)$ constraints. 
Note that we have to do this while keeping the total number of variables upper-bounded by some function in $k$.

For an easier understanding and the analysis of the algorithm we give the following definition.
\begin{definition}\label{def:segments}
	Let $U \subseteq V(G')$ be a set of vertices of interest and let $u,v \in U$.
	A path $P = (u=v_1,v_2, \dots, v_p = v)$  in graph $G'$, where all inner vertices are not in $U$, \ie $v_i \notin U$ for all $i \in \{ 2, 3, \dots, p-1\}$,
	is called a \emph{segment} from $u$ to $v$. We denote it as $S_{u,v}$.
\end{definition}
%From the definition it follows, that there are $O(k^2)$ segments in $G'$. \todo{True but does not follow directly just from the definition}
Note from \cref{def:segments} we get that $S_{u,v} \neq S_{v,u}$, since we consider paths to be directed. Observe that a temporal path in $G'$ between two vertices of interest is either a segment, or consists of a sequence of some segments. Furthermore, since we have at most $4k$ interesting vertices in $G'$, we can deduce the following important result.
\begin{corollary}\label{obs:FPT-k2segments}
	There are at most $O(k^2)$ segments in $G'$.
\end{corollary}
%In the analysis of the algorithm we rely heavily on the fact that there are $O(k^2)$ segments in $G'$.
%, which we prove later on (see~\cref{obs:FPT-k2segments}).
%For a pair $u,v$ of vertices in $U$, there are $O(k!)=O(k^k)$ possible paths in $G'$ between them.

%\subsubsection{Guessing necessary structures \label{sec:FPT-guessing}}
Once the sets $U, U^*$ and $Z^*$ are determined, we are ready to start guessing  the necessary structures.
Note that whenever we say that we guess the fastest temporal path between two vertices, we mean that we guess the underlying path of a representative fastest temporal path between those two vertices.
In the case when we want to guess the fastest path from $u$, that (somehow) passes through the vertex $x$, and goes directly to $v$ via an edge, we write it as a fastest path of the form $u \leadsto x \rightarrow v$.
%Similarly it holds if we want to include more vertices that have to be visited by the path. If there is an edge (\ie a unique fastest path) between two vertices, we denote it by $\rightarrow$,
%if the fastest path between two vertices is not uniquely determined, we denote it by $\leadsto$.
We guess the following structures.
%\noindent\textbf{Basic guesses:}
\begin{enumerate}[G-1.]
	\setcounter{enumi}{\value{guesscounter}}
	\item \label{FPT-guessFTPamongU}
	The fastest temporal paths between all pairs of vertices of $U$.
	For a pair $u,v$ of vertices in $U$, there are $k!$ possible paths in $G'$ between them. 
	Therefore, we have to try all $O(k^k)$ possible paths, where at least one of them will be a fastest temporal path from $u$ to $v$, respecting the values from $D$.
	%If this does not happen we conclude that $D$ is a \textsc{No} instance of \deltaExact.
	Repeating this procedure for all pairs of vertices $u,v \in U$ we get $O((k^k)^{k^2})=O(k^{k^3})$ different variations of the fastest temporal paths between all pairs of vertices in $U$.
	\item \label{FPT-guessFTPamongZstar}
	The fastest temporal paths between all pairs of vertices in $Z^*$, 
	which by similar arguing as for vertices in $U$, gives us $O(k^{k^3})$ guesses.
	\item \label{FPT-guessFTPamongUstar}
	The fastest temporal paths between all pairs of vertices in $U^*$.
	This gives us $O(k^{k^6})$ guesses.
	\item \label{FPT-guessFTPamongUandUstar}
	The fastest temporal paths from vertices of $U$ to vertices in $U^*$,
	and vice versa, the fastest temporal paths from vertices in $U^*$ to vertices in $U$.
	This gives us $O(k^{k^4})$ guesses.
	\item \label{FPT-guessFTPamongUandZstar}
	The fastest temporal paths from vertices of $U$ to vertices in $Z^*$,
	and vice versa.
	This gives us $O(k^{k^3})$ guesses.
	\item \label{FPT-guessFTPamongUstarandZstar}
	The fastest temporal paths from vertices of $U^*$ to vertices in $Z^*$,
	and vice versa.
	This gives us $O(k^{k^4})$ guesses.
	\setcounter{guesscounter}{\value{enumi}}
\begin{enumerate}[G-1.]
	\setcounter{enumi}{\value{guesscounter}}
	\item \label{FPT-guessFTPamongv2z2}
	\textbf{Inner segment guess I}.
	Let $S_{u,v} = (u=v_1,v_2, \dots, v_p = v)$ and $S_{w,z} = (w=z_1,z_2, \dots, z_r = z)$ be two segments,
	first one between vertices $u,v \in U$, second one between vertices $w, z \in U$.
	We want to guess the following fastest temporal path
	$v_2 \rightarrow u \leadsto w \rightarrow z_2$. 
	We repeat this procedure for all pairs of segments.
	Since there are $O(k^2)$ segments in $G'$,
	there are $O(k^{k^5})$ possible paths of this form. \\
	Recall that $S_{u,v}\neq S_{v,u}$ for every $u,v\in U$. Furthermore note that we did not assume that the sets $\{u,v\}$ and $\{w,z\}$ are disjoint. Therefore, by repeatedly making the above guesses, we also guess the following fastest temporal paths: 
	${v_2 \rightarrow u \leadsto z \rightarrow z_{r-1}}$,\ \ \ 
	${v_2 \rightarrow u \leadsto v \rightarrow v_{p-1}}$,\ \ \  
	${v_{p-1} \rightarrow v \leadsto w \rightarrow z_{2}}$,\ \ \  
	${v_{p-1} \rightarrow v \leadsto z \rightarrow z_{r-1}}$, and  
	${v_{p-1} \rightarrow v \leadsto u \rightarrow v_{2}}$.
	%
	For an example see~\cref{fig:FPT-guessG4}.
	\item \label{FPT:guess-uToSegmentz2}
	\textbf{Inner segment guess II}.
	Let $S_{u,v} = (u=v_1,v_2, \dots, v_p = v)$ be a line segment in $G'$,
	and let $w \in U \cup Z^*$ be either a vertex of interest or a representative vertex of a tree, whose clipped vertex is a vertex of interest.
	We want to
	guess the following fastest temporal paths
	$w \leadsto u \rightarrow v_2$, $w \leadsto v \rightarrow v_{p-1} \rightarrow \cdots \rightarrow v_2$,
	$v_2 \rightarrow u \leadsto w$, $v_2 \rightarrow v_3 \rightarrow \cdots v \leadsto w$.
	\\
	For a fixed segment $S_{u,v}$ and a fixed vertex $w \in U \cup Z^*$ we have $O(k^k)$ different possible such paths, therefore we make $O(k^{k^4})$ guesses for these paths.
	%
	For an example see~\cref{fig:FPT-guessG5}.
	%All together we made $O(k^{k^4}$  new guesses for the structure of all of the above mentioned fastest temporal paths.
	%Each of the previous partial ILP instances is therefore extended into $O(k^6)$ new ILP instances.
	\item \label{FPT:guess-splitFromAnotherSegmentAndPaths}
	\textbf{Split vertex guess I}.
	Let $S_{u,v} = (u=v_1,v_2, \dots, v_p = v)$ be a line segment in $G'$, and let us
	fix a vertex $v_i \in S_{u,v} \setminus \{u,v\}$.
	In the case when $S_{u,v}$ is of length $4$, the fixed vertex $v_i$ is the middle vertex (\ie vertex that is on distance two from $u$ and $v$ in $S_{u,v}$).
	If $S_{u,v}$ is not of length~$4$, we fix an arbitrary vertex $v_i \in S_{u,v} \setminus \{u,v\}$.
	Let 
	$S_{w,z} = (w=z_1,z_2, \dots, z_r = z)$ be another segment in $G'$.
	Note that all inner vertices of $S_{w,z}$ (\ie vertices in $S_{w,z} \setminus \{w,z\}$)
	can be reached from $v_i$ by some path that has to go through $w$ or $z$.
	Therefore, the values of the duration of the fastest path from $v_i$ to vertices on 
	the segment
	$S_{w,z}$ have to increase up to a certain point, when traversing the segment coming through $w$ and coming through $z$.
	We split the analysis into two cases.
	\begin{enumerate}
		\item 
		There is a single vertex $z_j \in S_{w,z}$ for which the duration from $v_i$ is the biggest.
		More specifically, $z_j \in S_{w,z} \setminus \{w,z\}$ is the vertex with the biggest value  $D_{v_i,z_j}$.
		We call this vertex a \emph{split vertex of $v_i$ in the segment $S_{wz}$}.
		Then it holds that $D_{v_i,z_2} < D_{v_i,z_3} < \dots < D_{v_i,z_j}$ and 
		$D_{v_i,z_{r-1}} < D_{v_i,z_{r-2}} < \dots < D_{v_i,z_j}$.
		From this it follows that the fastest temporal paths from $v_i$ to $z_2, z_3, \dots, z_{j-1}$ go through $w$,
		the fastest temporal paths from $v_i$ to $z_{r-1}, z_{r-2}, \dots, z_{j+1}$ go through $z$.
		We now want to guess which vertex $w$ or $z$ is on a fastest temporal path from $v_i$ to $z_j$.
		Similarly,
		all fastest temporal paths starting at $v_i$ have to go either through $u$ or through $v$,
		which also gives us two extra guesses for the fastest temporal path from $v_i$ to $z_j$.
		Therefore, all together we have $4$ possibilities on how the fastest temporal path from $v_i$ to $z_j$ starts and ends.
		Besides that we want to guess also how the fastest temporal paths from $v_i$ to $z_{j-1}, z_{j+1}$ start and end.
		Note that one of these is the subpath of the fastest temporal path from $v_i$ to $z_j$, and the ending part is uniquely determined for both of them,
		\ie to reach $z_{j-1}$ the fastest temporal path travels through $w$, and to reach $z_{j+1}$ the fastest temporal path travels through $z$.
		Therefore we have to determine only how the path starts, namely if it travels through $u$ or $v$.
		This introduces two extra guesses.
		For a fixed $S_{u,v}, v_i$ and $S_{w,z}$ we find the vertex $z_j$ in polynomial time, 
		or determine that $z_j$ does not exist.
		We then make four guesses where we determine how the fastest temporal path from $v_i$ to $z_j$ passes through vertices $u,v$ and $w,z$ and 
		for each of them two extra guesses to determine the fastest temporal path from $v_i$ to $z_{j-1}$ and from $v_i$ to $z_{j+1}$.
		We repeat this procedure for all pairs of segments,
		which results in producing $O(k^{k^6})$ new guesses.
		Note, $v_i \in S_{u,v}$ is fixed when calculating the split vertex for all other segments $S_{w,z}$.
		\item 
		There are two vertices $z_j, z_{j+1} \in S_{w,z}$ for which the duration from $v_i$ is the biggest.
		More specifically, $z_j, z_{j+1} \in S_{w,z} \setminus \{w,z\}$ are the vertices with the biggest value  $D_{v_i,z_j} = D_{v_i,z_{j+1}}$.
		Then it holds that $D_{v_i,z_2} < D_{v_i,z_3} < \dots < D_{v_i,z_j} = D_{v_i,z_{j+1}} > D_{v_i,z_{j + 2}} > \cdots > D_{v_i,z_{r-1}}$.
		From this it follows that the fastest temporal paths from $v_i$ to $z_2, z_3, \dots, z_{j}$ go through $w$,
		the fastest temporal paths from $v_i$ to $z_{r-1}, z_{r-2}, \dots, z_{j+1}$ go through $z$.
		In this case we only need to guess the following two fastest temporal paths 
		$u \leadsto w \rightarrow z_2$
		and $u \leadsto z \rightarrow z_{r-1}$.
		Each of this paths we then uniquely extend along the segment $S_{w,z}$ up to the vertex $v_j$, resp.~$v_{j+1}$,
		which give us fastest temporal paths from $u$ to $v_j$ and from $u$ to $v_{j+1}$.
		Note that in this case we do not introduce any new guesses, as we have aready guessed 
		the fastest paths of the form
		$u \leadsto w \rightarrow z_2$
		and $u \leadsto z \rightarrow z_{r-1}$
		(see guess~G-\ref{FPT:guess-uToSegmentz2}).
	Note that this case results also in knowing the fastest paths from the vertex $v_i \in S_{u,v}$ to $w,z \in S_{w,z}$ for all segments $S_{w,z}$, 
	\ie we know the fastest paths from a fixed $v_i \in S_{u,v}$ to all vertices of interest in $U$.
	%
	For an example see~\cref{fig:FPT-guessG6}.
	\item \label{FPT:guess-splitFromUtoAnotherSegment}
	\textbf{Split vertex guess II}.
	Let $w \in U \cup Z^*$ be either a vertex of interest or a representative vertex of a tree, whose clipped vertex is a vertex of interest,
	and let $S_{u,v} = (u=v_1,v_2, \dots, v_p = v)$ be a line segment in $G'$.
	We want to guess a split vertex of $w$ in $S_{u,v}$, and the fastest temporal path that reaches it.
	We again have two cases,
	first one where $v_i$ is a unique vertex in $S_{u,v}$ that is furthest away from $w$,
	and the
	second one where $v_i, v_{i+1}$ are two incident vertices in $S_{u,v}$, that are furthest away from $w$.

@@

	In first case we know exactly how the fastest paths from $w$ to all vertices $v_j \in S_{u,v} \setminus \{v_i\}$
	travel through the segment $S_{u,v}$ (\ie either through $u$ or $v$).
	Therefore we have to guess how the fastest path from $w$ reaches vertex $v_i$,
	we have two options, either it travels through $u \rightarrow v_2 \rightarrow \cdots \rightarrow v_{i-1} \rightarrow v_i$
